Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-12 paper/paper.tex
More typos

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
much of brain activity is dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing tentative evidence that
variation of the grey-matter volume in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally,
we derive explicit hypotheses that could be tested in targeted neuroscience
experiments in the future
and
we detail how our process model relates
to previous cognitive and theoretical accounts of DMN function.

@@

  \subsubsection{Self-training and the loss function}
  \label{sec:self}
Successful learning in brains and computer algorithms may not be possible without a
defined optimization goal \textemdash the \textit{loss function}.
The action $a$ chosen in state $s$ according to the policy matrix defined in
\eqref{eq:policy} yields a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
One such cycle yields a new \textit{experience} $e = (s,a,r,s')$.
Each cycle represents a behavior unit of the agent
and is recorded in replay memory buffer \textemdash
which we hypothesize to involve especially the
HC \textemdash, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1}
+ \delta \theta_{k}$ of the parameters for its approximate model of the
$Q$-value function. Step-by-step model parameter updates warrant a learning process and definition of a loss
function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss
function \eqref{eq:oracle} as we outline in the following.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Experience replay consists in sampling
% (uniform or importance-weighted
% \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
batches of experiences $e$
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the
Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the estimation of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective, experience replay can be manifested as the
re-occurrence of neuron spiking sequences that have also been measured during
specific prior actions or environmental states. The HC is a strong candidate
for contributing to such neural reinstantiation of behavioral episodes as
neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
Importantly, neural encoding of abstract representations of space and meaning 
may extent to several parts of the DMN \citep{behrens_grid}.
Refer to
Fig. \ref{fig:rl_process_chart}.

>>>>

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
much of brain activity is dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing tentative evidence that
variation of the grey-matter volume in DMN regions
is linked to the reward circuitry (Fig. \ref{fig:dmn}),
thus linking two literatures with currently scarce cross-references.
Finally,
we derive explicit hypotheses that could be tested in targeted neuroscience
experiments in the future
and
we detail how our process model relates
to previous cognitive and theoretical accounts of DMN function.

@@

  \subsubsection{Self-training and the loss function}
  \label{sec:self}
Successful learning in brains and computer algorithms may not be possible without a
defined optimization goal \textemdash the \textit{loss function}.
The action $a$ chosen in state $s$ according to the policy matrix defined in
\eqref{eq:policy} yields a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
One such cycle yields a new \textit{experience} $e = (s,a,r,s')$.
Each cycle represents a behavior unit of the agent
and is recorded in replay memory buffer \textemdash
which we hypothesize to involve especially the
HC \textemdash, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1}
+ \delta \theta_{k}$ of the parameters for its approximate model of the
$Q$-value function. Step-by-step model parameter updates warrant a learning process and definition of a loss
function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss
function \eqref{eq:oracle} as we outline in the following.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Experience replay consists in sampling
% (uniform or importance-weighted
% \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
batches of experiences $e$
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the
Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the estimation of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective, experience replay can be manifested as the
re-occurrence of neuron spiking sequences that have also been measured during
specific prior actions or environmental states. The HC is a strong candidate
for contributing to such neural reinstantiation of behavioral episodes as
neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
Importantly, neural encoding of abstract representations of space and meaning 
may extent to several parts of the DMN \citep{behrens_grid} (see
Fig. \ref{fig:rl_process_chart}).
